### 2014-03-31 paper/abstract.tex
paper work, adding tovi comments

\begin{abstract}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for adding interactivity by removing interior material to form pipes and cavities. These can be used to sense input and generate haptic as well as visible output. We describe the design space of pipes and hollow chambers for interaction design: a variety of openings, topologies, and inserted media for pipes create diverse inputs and outputs.  We present a technique and design tool for routing pipes of various topologies through the interior of 3D printed parts.  Our design tool is integrated into a 3D model manipulation program.  We use two distinct routing algorithms.  One, for creating particular points of interaction or integrating with existing electronics on the model's surface, allows users to select begin and end points for pipes, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  The second, for creating, e.g., novel neon signs, has users supply a desired internal shape which we fit a pipe route to: for this we developed a novel routing algorithm and offer proof of key traits.  We present several totally tubular prototypes we created using this tool to show its flexibility and potential, as well as to explore new points in the pipe design space.
\end{abstract}

>>>>

\begin{abstract}
\tovi{can be condensed}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for adding interactivity by removing interior material to form pipes and cavities. We describe the design space of pipes and hollow chambers for interaction design: a variety of openings, topologies, and inserted media for pipes create diverse inputs and outputs.  We present a technique and design tool for routing pipes of various topologies through the interior of 3D printed parts.  Our design tool is integrated into a 3D model manipulation program.  We use two distinct routing algorithms.  One, for creating particular points of interaction or integrating with existing electronics on the model's surface, allows users to select begin and end points for pipes, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  The second, for creating, e.g., novel neon signs, has users supply a desired internal shape which we fit a pipe route to: for this we developed a novel routing algorithm and offer proof of key traits.  We present several totally tubular prototypes we created using this tool to show its flexibility and potential, as well as to explore new points in the pipe design space.
\end{abstract}

### 2014-03-31 paper/designspace.tex
paper work, adding tovi comments

Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Since we are interested in interactive devices, we distinguish a ``user side" - part of the surface or interior of an object facing its user; and a ``system side" - part of the object (either interior or exterior) where other hardware such as electronics will be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' or ``open on the system side'' to have its contents \emph{physically accessible} to an interacting user or the system \emph{at interaction time}.  A pipe which is ``open on the user side'' or ``closed on the system side'' is physically cut off from the user or machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/types.png}
\caption{Types of pipes.  A shows a pair of open tubes filled with copper paint to power an LED.  B shows a return pipe with EL wire threaded through it.  C shows two semi-closed pipes with rubber membranes at their ends.  In D, a rabbit has a fully enclosed chamber to modify its weight.}
\label{fig:pipespace}
\end{figure}

@@

\subsection{Design of Pipes}

Pipe designs may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe, parietal lobes, etc.  An interface focused on the internal path of the pipe is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the pipe.

>>>>

\tovi{this seems like a forced and
unnecessary definition. Do we really
need to distibguish the two ends?}
Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Since we are interested in interactive devices, we distinguish a ``user side" - part of the surface or interior of an object facing its user; and a ``system side" - part of the object (either interior or exterior) where other hardware such as electronics will be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' or ``open on the system side'' to have its contents \emph{physically accessible} to an interacting user or the system \emph{at interaction time}.  A pipe which is ``open on the user side'' or ``closed on the system side'' is physically cut off from the user or machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/types.png}
\caption{Types of pipes.  A shows a pair of open tubes filled with copper paint to power an LED.  B shows a return pipe with EL wire threaded through it.  C shows two semi-closed pipes with rubber membranes at their ends.  In D, a rabbit has a fully enclosed chamber to modify its weight.}
\label{fig:pipespace}
\end{figure}

@@

\subsection{Design of Pipes}
\tovi{This definition is a little bit vague.}
Pipe designs may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe, parietal lobes, etc.  An interface focused on the internal path of the pipe is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the pipe.

### 2014-03-31 paper/exampleobjects.tex
paper work, adding tovi comments

\valkyrie{one way of combatting the enormousness of this paper is to put all these example objects figures into a single figure* instead}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/brain.jpg}
\caption{A touch-sensitive brain whose tubes are filled with conductive paint.  Sensing is done on a single wire via SFCS.  Inset shows the internal structure of the tubes generated by our design tool.}
\label{fig:toys}
\end{figure}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/bunny.jpg}
\caption{This rabbit has an open tube leading to its mouth and a semi-closed tube capped with flexible material in its back.  When air is sucked in through the open tube, the back becomes stiff as though the rabbit has breathed in.  Inset shows the interior structure.}
\label{fig:breathe}
\end{figure}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/radio.png}
\caption{This radio is assembled from traditional electronic components connected by copper-filled tubes.  The case was designed to allow the components to recess into it slightly.  Inset shows the internal structure of the tubes generated by our design tool.}
\label{fig:radio}
\end{figure}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/pens.jpg}
\caption{This pen holder a modified FlyEye technique to sense the presence or absence of an object in each of its tubes.  Inset shows the internal structure of the tubes generated by our design tool.}
\label{fig:pens}
\end{figure}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/maze.jpg}
\caption{Our maze game has a large single particle trapped inside a fully enclosed tube.  It offers hours of fun!}
\label{fig:maze}
\end{figure}

>>>>

\begin{figure*}[h!]
\centering
    \includegraphics[width=7in]{figures/examples.png}
\caption{A series of example objects created using our system.  A) shows a portable radio.  B) shows a presence-aware pen holder.  C) is a rabbit which ``breathes'' using wind pipes we built.  D) shows a maze game we created.  E) is a touch-sensitive brain toy.  \valkyrie{finalize order of example objects.  what order is best?}}
\label{fig:examples}
\end{figure*}

@@

@@

@@

@@

@@

### 2014-03-31 paper/tubes.tex
paper work, adding tovi comments

\subsubsection{Routing}
In graph theory, a semi-Eulerian graph is one which has exactly two edges of odd degree.  This kind of graph supports an Euler tour: a path that touches every edge exactly once, beginning and ending at the two nodes of odd degree (in this case, the user-selected start and end nodes).  To allow for a single path through a pipe for post-print media insertion, we need to find a semi-Eulerian graph for which a user's input graphics are a subset, and subsequently find an Euler tour on that graph.

@@

>>>>

\subsubsection{Routing}
\tovi{Still too long....}
In graph theory, a semi-Eulerian graph is one which has exactly two edges of odd degree.  This kind of graph supports an Euler tour: a path that touches every edge exactly once, beginning and ending at the two nodes of odd degree (in this case, the user-selected start and end nodes).  To allow for a single path through a pipe for post-print media insertion, we need to find a semi-Eulerian graph for which a user's input graphics are a subset, and subsequently find an Euler tour on that graph.

@@

\subsection{Fabrication Techniques}

The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  The two machines we used to create our example objects were the Objet Connex 260 and the Makerbot Replicator 2.

The Objet is an inkjet-based 3D printer which jets drops of polymer in very fine layers, curing each layer with an ultraviolet lamp.  It offers a wide range of material types, including clear/opaque, flexible/hard, and black/white.  Additionally, it permits the mixing of two different materials in a single print: a mixed material could include 70\% flexible and 30\% hard, for example, to be firm but somewhat flexible.  The flexible materials can be inflated when they are thin: they elongate 200\% before breaking.  We have found $.05$ inches to be a good thickness for inflatable materials.  Although this machine has many extraordinary capabilities, it unfortunately cannot print overhangs of greater than $14^{\circ}$ without support material.  Due to this, support avoidance is difficult; our main tactic for removing support material has been to make the \emph{openings} to pipes as large as possible.  This allows us to flush the material most easily.  The support structure used by this machine is a loose matrix of material that can be removed by hand from the surface, blasted out with a water jet from cavities, or dissolved in lye (the enclosing model material does not dissolve).

The Makerbot Replicator 2 is a fused-deposition modeling (FDM) printer, in which we use exclusively Polylactic Acid (PLA) bioplastics.  More material properties are being explored in PLA: there are now transparent and flexible materials.  They cannot be blended like those in the Objet.  The Makerbot has a fairly coarse resolution compared to the Objet, however it permits overhangs up to $45^{\circ}$ without support and also allows bridging (short spans of material that bridge the gap between two structures with no support underneath).  Although spheres and sideways cylinders have a $>45^{\circ}$ overhang in some places, they are also printable without support.  Due to the FDM process it can even recover from greater overhangs and longer bridges, as the material builds up and can reorganize into a structure \valkyrie{how do I say this?  do I need a photo?  print recovery is pretty awesome-looking.}.  The support used by the Makerbot is made of the same material as the model itself, just printed in a different pattern and attached over as small a surface area as possible to ease removal.  Support inside of a pipe may require pliers, or be infeasible to remove dependent upon geometry.  Meshmixer has a built-in tool for minimizing support for prints; it re-orients the part to keep as much surface area as possible within the 45 degree constraint.  In future work we would like to explore a tool in which a user can select surfaces to prioritize support avoidance for.

Assembly is easiest on both machines when the model does not require support material.  For especially complex geometries, models can be ``cut up'' \bjoern{This negates some of the benefits compared to other manufacturing techniques.} into assembleable pieces that evade support requirements or ease material removal.  We believe that an automatic tool for this task would be highly beneficial, however it is outside the scope of this work.
